Federated Learning's Blessing:\\ FedAvg has Linear Speedup

### 2020-06-05 source/sec_intro.tex
Update on Overleaf.

% Our contributions
% \subsection{Contributions}
\subsection{Our Contributions.}
We provide a comprehensive convergence analysis
of FedAvg and its accelerated variants in convex stochastic problems with both system and statistical heterogeneity (see Table~\ref{tb:convergencerateintro}).
Our results demonstrate the desirable linear speedup property with respect to the number of participating clients.
More specifically, for strongly convex and smooth problems, we show 
that FedAvg admits a convergence rate of 
{\small{$\cO(1/NT)$}}, and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}. For convex and smooth objectives, the rate is {\small{$\cO(1/\sqrt{NT})$}}. 
Moreover, we show that FedAvg with Nesterov acceleration also enjoys the same rates and speedup. 
%The same conclusions are equally applied for Nesterov accelerated FedAvg. 
To the best of our knowledge, this is the first linear speedup result for FedAvg and accelerated variants in the fully heterogeneous FL setting. Furthermore, 
we propose a new variant of FedAvg and provide the first improved convergence rate for stochastic momentum-accelerated FedAvg in the popular overparamterized setting.
Last but not least, we systematically examine the convergence rate of
FedAvg and its accelerated variants on both a synthetic dataset and  a real-world dataset. The empirical observations are well-aligned with the obtained theoretical findings.

\begin{table}[h!]
\centering
{\small
\begin{tabular}{|c|c|c|c|}\hline 
	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex  & Overparameterized \\ \hline \hline
	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$   & $\cO(\exp(-\frac{NT}{E}))$    \\ \hline
	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$ & $\cO(\exp(-\frac{KT}{E}))$     \\ \hline
\end{tabular}
}
\caption{Our convergence results for FedAvg and Nesterov accelerated FedAvg in this paper.}
\label{tb:convergencerateintro}
\end{table}

>>>>

% Our contributions
% \subsection{Contributions}
\subsection{Our Contributions.}
We provide a comprehensive convergence analysis
of FedAvg and its accelerated variants in the presence of both system and statistical heterogeneity (see Table~\ref{tb:convergencerateintro}).
Our results demonstrate the desirable linear speedup property with respect to the number of participating clients.
More specifically, for strongly convex and smooth problems, we show 
that FedAvg admits a convergence rate of 
{\small{$\cO(1/NT)$}}, and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}. For convex and smooth objectives, the rate is {\small{$\cO(1/\sqrt{NT})$}}. 
Moreover, we show that FedAvg with Nesterov acceleration also enjoys the same rates and speedup. 
%The same conclusions are equally applied for Nesterov accelerated FedAvg. 
To the best of our knowledge, this is the first linear speedup result for FedAvg and accelerated variants in the fully heterogeneous FL setting. Furthermore, 
we propose a new variant of FedAvg and provide the first improved convergence rate for stochastic momentum-accelerated FedAvg in the popular overparamterized setting.
Last but not least, we systematically examine the convergence rate of
FedAvg and its accelerated variants on both a synthetic dataset and  a real-world dataset. The empirical observations are well-aligned with the obtained theoretical findings.

\begin{table}[h!]
\centering
{\small
\begin{tabular}{|c|c|c|c|}\hline 
	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex  & Overparameterized \\ \hline \hline
	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$   & $\cO(\exp(-\frac{NT}{E}))$    \\ \hline
	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$ & $\cO(\exp(-\frac{KT}{E}))$     \\ \hline
\end{tabular}
}
\caption{Our convergence results for FedAvg and accelerated FedAvg in this paper.}
\label{tb:convergencerateintro}
\end{table}
